---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# satpt

<!-- badges: start -->
<!-- badges: end -->

The goal of satpt is to identify the saturation point of multinomial responses from a survey using standard errors of sample proportions

## Installation

You can install the development version of `satpt` like so:

```{r, eval = FALSE}
#install.packages("remotes")
remotes::install_github("deboonstra/satpt")
```

## Usage

### Basic usage
For basic usage of `satpt` simply specify the responses of the survey in `y` and the data collection period information in `by`, as seen below. 
```{r, eval = FALSE}
data(ein)
satpt::satpt(y = ein$responses, by = ein$wave)
```

```{r, echo = FALSE}
data(ein)
satpt::satpt(y = ein$responses, by = ein$wave)
```

### Dimension names specification
When specification of the dimension names of the contigency table, sample proportions, and standard error is wanted, use one of these two methods.

#### Method 1
```{r, eval = FALSE}
res <- satpt::satpt(
  y = ein$responses,
  by = ein$wave,
  dimnames = c("Responses to survey", "Collection period")
)
res$counts # accessing contingency table
```

```{r, echo = FALSE}
res <- satpt::satpt(
  y = ein$responses,
  by = ein$wave,
  dimnames = c("Responses to survey", "Collection period")
)
res$counts # accessing contingency table
```

#### Method 2
```{r, eval = FALSE}
res <- satpt::satpt(
  y = ein$responses,
  by = ein$wave,
  dimnames = c("by" = "Collection period", "y" = "Responses to survey")
)
res$phat # accessing sample proportions table
```

```{r, echo = FALSE}
res <- satpt::satpt(
  y = ein$responses,
  by = ein$wave,
  dimnames = c("by" = "Collection period", "y" = "Responses to survey")
)
res$phat # accessing sample proportions table
```