---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# satpt

<!-- badges: start -->
[![R-CMD-check](https://github.com/deboonstra/satpt/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/deboonstra/satpt/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

This *R* package identifies the saturation point of multinomial responses from a survey using standard errors of sample proportions

## Installation
You can install the current released version of `satpt` from [CRAN](https://cran.r-project.org) with

```{r, eval = FALSE}
install.packages("satpt")
```

### Development version
To get a bug fix or to use a feature from the development version, you can install the development version of `satpt` from [GitHub](https://github.com/deboonstra/satpt).

```{r, eval = FALSE}
# install.packages("remotes")
remotes::install_github("deboonstra/satpt")
```

## Usage

For basic usage of `satpt` simply specify the responses of the survey in `y` and the data collection period information in `by`, as seen below. 
```{r, eval = FALSE}
library(satpt) # load package
data(ein) # load example ein data
res <- satpt::satpt(y = ein$q2, by = ein$wave)
print(res)
```

```{r, echo = FALSE}
data(ein)
res <- satpt::satpt(y = ein$q2, by = ein$wave)
print(res)
```

Saturation of each individual response category may be examined graphically while comparing the standard errors to the saturation threshold.

```{r, eval = FALSE}
graphics::par(oma = c(0, 0, 0, 8))
plot(res)
# adding legend
satpt::legend_right(
  legend = "Saturation\nthreshold",
  col = "firebrick", lty = 3, lwd = 2,
  cex = 0.75
)
```

```{r, echo = FALSE}
graphics::par(oma = c(0, 0, 0, 8))
plot(res)
# adding legend
satpt::legend_right(
  legend = "Saturation\nthreshold",
  col = "firebrick", lty = 3, lwd = 2,
  cex = 0.75
)
```

# Methodology

A paper explaining the methodology of this package is currently being written. When it is published, a link to that paper will be provided here. 
